Revisiting the Classics: Online RL in the Programmable Data-Plane

### 2021-02-05 paper/paper.tex
Reward packet gen.
Preliminary fixes to signalling tests, paper expansion

\xpatchbibmacro{name:andothers}{%
	\bibstring{andothers}%
}{%
	\bibstring[\emph]{andothers}%
}{}{}

@@

@@

% DOI
\acmDOI{}

% ISBN
\acmISBN{}

%Conference
%\acmConference[Submitted for review to SIGCOMM]{}
%\acmYear{2018}
%\copyrightyear{}

@@

%	\author{Paper \# XXX, XXX pages}
	 \author{Kyle A. Simpson}
	 \orcid{0000-0001-8068-9909}
	 \affiliation{%
	   \institution{University of Glasgow}
	   \city{Glasgow} 
	   \state{Scotland}
	 }
	 \email{k.simpson.1@research.gla.ac.uk}
	 \author{Dimitrios P. Pezaros}
\orcid{0000-0003-0939-378X}
	 \affiliation{%
	\institution{University of Glasgow}
	\city{Glasgow} 
	\state{Scotland}
}
\email{Dimitrios.Pezaros@gla.ac.uk}

@@

\section{Motivation}

\begin{figure}
	\begin{subfigure}{0.45\linewidth}
		\resizebox{\linewidth}{!}{
			\begin{tikzpicture}
					\node[circle, draw] (state) {$S$};
					\node[circle, draw] (state') at ($(state) + (0, -2)$) {$S'$};

@@

					\draw[dotted, ->, bend left = 30] (state) -- (agent);
					\draw[->] (agent) -- (action);
					\draw[dotted, ->] (action) -- (state);
		\end{tikzpicture}}
		\caption{Theory: state measurement, action computation, and learning are zero-cost.}
	\end{subfigure}
	\begin{subfigure}{0.45\linewidth}
		\resizebox{\linewidth}{!}{
			\begin{tikzpicture}
					\node[circle, draw] (state) {$S$};
					\node[circle, draw] (state') at ($(state) + (0, -1.5)$) {$S'$};
					\node[circle, draw] (state'') at ($(state') + (0, -1.5)$) {$S''$};

@@

What is our goal?

\begin{itemize}
	\item Research into asynchronous RL shows that clever reordering and mitigation can reduce the effects of state slippage on learning time and on accuracy of actions. However, it cannot completely eliminate the effect of state evolving due to the costs of action computation, execution, and policy updates~\cite{DBLP:journals/firai/TravnikMSP18}.

	\item What if we do stop the world (i.e., block operation) to ensure this doesn't happen? In case of congestion control, significant throughput drops (as DNN inference takes \SI{30}{\milli\second}), so asynchrony is necessary~\cite{DBLP:journals/corr/abs-1910-04054}.

	\item Training deep RL models has its own costs (need many parallel simulations). What if we only have one stream of experience to learn from? E.g., evolution of the system is too difficult to predict and model. In general, DNNs can't be trained effectively from a single, online, real-time stream of experience.

	\item PDP hardware is advertised as supporting P4, EBPF... But almost all implementations compile to a more complex/expressive language. Presence of externs is almost universal when it comes to in-switch compute---i.e., Tofino Native Architecture\footnote{\url{    https://www.barefootnetworks.com/products/brief-p4-studio/}}, Netronome Micro-C\footnote{\url{https://www.netronome.com/products/datapath-programming-tools/}}, P4$\rightarrow$NetFPGA toolchain\footnote{\url{https://github.com/NetFPGA/P4-NetFPGA-public}}.
\end{itemize}

What aspects of the state-of-the-art are relevant here?

\begin{itemize}
	\item Hmmm...
\end{itemize}

@@

\begin{figure}
	\centering
	\resizebox{1.2\linewidth}{!}{\begin{tikzpicture}
		\node[fill=white!80!uofgcobalt, draw=black, minimum height=1cm, minimum width=2.9cm](p4-master){};
		\node[below right, inner sep=2pt] at (p4-master.north west) {\small P4 Master Island};
		\node[fill=white!90!uofgcobalt, draw, rectangle, rounded corners=0.05cm, above=0.1] (p4-rules) at (p4-master.south) {\begin{varwidth}{2.5 cm}\scriptsize \centering Flow Control Rules\end{varwidth}};

		\node[fill=white!80!uofgcobalt, draw=black, minimum height=1cm, minimum width=2cm, below = 0.1 of p4-master](sw1){};
		\node[fill=white!80!uofgcobalt, draw=black, minimum height=1cm, minimum width=2cm, below right= 0.1 of sw1.north west](sw2){};
		\node[fill=white!80!uofgcobalt, draw=black, minimum height=1cm, minimum width=2cm, below right= 0.1 of sw2.north west](switch){};
		\node[below right, inner sep=2pt] at (switch.north west) {\small P4 Island};
		\node[fill=white!90!uofgcobalt, draw, rectangle, rounded corners=0.05cm, above=0.1] (oswlc) at (switch.south) {\begin{varwidth}{1.5 cm}\scriptsize \centering Control Packet Match\end{varwidth}};

		\node[fill=white!80!uofgcobalt, draw=black, minimum height=1.5cm, minimum width=2.9cm, below= 0.1 of switch](others){};
		\node[below right, inner sep=2pt] at (others.north west) {\small Other Islands};
		\node[fill=white!90!uofgcobalt, draw, rectangle, rounded corners=0.05cm, above=0.1] (others-etc) at (others.south) {\begin{varwidth}{2.5 cm}\scriptsize \centering \textellipsis \end{varwidth}};
		\node[fill=white!90!uofgcobalt, draw, rectangle, rounded corners=0.05cm, above=0.1] (others-int) at (others-etc.north) {\begin{varwidth}{2.5 cm}\scriptsize \centering In-band Network Telemetry\end{varwidth}};

		\node[fill=white!80!uofgsunshine, draw=black, minimum height=5cm, minimum width=5cm](rli) at ($(p4-master.east) + (3, -2)$){};
		\node[below right, inner sep=2pt] at (rli.north west) (rli-text) {\small RL Island};

		\node[draw=black, dotted, minimum height=1cm, minimum width=2cm](wq-out) at ($(rli-text.south) + (0.5,-0.5)$){};
		\node[below right, inner sep=2pt] at (wq-out.north west) {\small CLS};
		\node[fill=white!90!uofgsunshine, draw, rectangle, rounded corners=0.05cm, above=0.1] (wq-in) at (wq-out.south) {\begin{varwidth}{2.5 cm}\scriptsize \centering Work Queue\end{varwidth}};

@@

		\draw[->] (loop.north) [out=90,in=45,looseness=0.75] to node[pos=0.5, rotate=0, anchor=center, above] {Rules, Tables}(p4-rules.north east);

		\draw[->] (p4-rules.west)  to [out=180,in=180,looseness=2]  node[pos=0.5, rotate=90, anchor=center, above] {Rules}(sw1);
		\draw[->] (p4-rules.west)  to [out=180,in=180,looseness=2] (sw2);
		\draw[->] (p4-rules.west)  to [out=180,in=180,looseness=2] (switch);

		\draw[->] (oswlc.east)  to node[midway, above, sloped] {\begin{varwidth}{2.5cm}\tiny \centering Control, Requests,\\State\end{varwidth}} (wq-in.west);
		\draw[->] (others-int.east)  to node[midway, above, sloped] {\tiny State} (wq-in.west);
		\draw[<->, dotted] (wq-in.east)  to (loop.west);
		\draw[->, dotted] (loop.south) to (policy-cls-out);

@@

		\draw[->] (loop) to node[midway, above, sloped] {\tiny NNR} (po-me-1);
		\draw[->] (po-me-1) to node[midway, above, sloped] {\tiny NNR} (po-me-2);
		\draw[->] (po-me-2) to node[midway, above, sloped] {\tiny NNR} (po-me-3);
		\draw[->] (po-me-3) to [out=-45,in=0,looseness=2.15] node[midway, above, sloped] {\tiny NNR} (loop);
		\end{tikzpicture}}
	\caption{Architecture for generalised RL agent on Netronome hardware.\label{fig:netro-arch}}
\end{figure}

@@

			\item[$t_3$]
			\begin{description}
				\item[PDP] Modify local tables across ME/Island boundaries (same device), verify rule present (or get ACK).
				\item[VNF] Generate OpenFlow control messages, send OF message to switch, verify rule present (or get ACK).
			\end{description}
		\end{description}
		\item HYPOTHESIS: PDP will present a reduction in these three times for per-flow state.
		\item QUESTION: Does same apply to learning?
		\item NOTE: since we're not actually assessing the \emph{quality} of output decisions, we can cut corners on what data is sent out and where. I.e., in-band network telemetry island can be skipped, if we move over representative volumes of state to/from the RL island with approximately correct delay as recorded by other studies.
	\end{enumerate}
\end{itemize}
I can think of one way to evaluate this if we employ classification performance:
\begin{enumerate}
	\item Run the existing work~\cite{DBLP:journals/tnsm/SimpsonRP20}, acquire a policy for one node after full info sharing.
	\item Load this policy onto the RL Island (\cref{fig:netro-arch}).
	\item Either feed in raw state packets, or feed in traffic if we can get INT working.
	\item For each level of quantisation needed, examine average performance.
	\item Don't compare against old numbers for final perf: use a ``VNF'' setup similar to above---this will probably suffer a bit compared to old results, where everything was on the same machine. Make it relative/normalised.
\end{enumerate}

\newpage

\section{Bullets on (intended) design.}
See also: \cref{fig:netro-arch}.
\begin{itemize}
	\item High-level overview.
	\begin{itemize}
		\item RL operations \emph{do not happen on the core path}.
		\item Another island (or set of cores) is dedicated to handling inputs, configuration, computation.
		\item Cross-island communication happens over \emph{rings} (Netronome-specific).
		\begin{itemize}
			\item E.g., this is how the P4 firmware propagates rules from master island to P4 workers.
		\end{itemize}
		\item Controller is \emph{always} co-hosted, interfaces over PCI-e. This also applies to Barefoot Tofino.
	\end{itemize}
	\item Policy configuration.
	\begin{itemize}
		\item (Re)-configurable at runtime.
		\item Custom DSCP marker acts as indicator of RL packets (i.e., bother with parsing past UDP).
		\item P4 parser extracts headers, parser state and payload passed over ring to another core dedicated to RL operations.
		\item Policy dimensions, size, action count, tiling strategies, learning parameters ($\alpha, \gamma, \epsilon$, related falloffs) can all be chosen and changed live.
	\end{itemize}
	\item State Handling.
	\begin{itemize}
		\item State input to model \emph{always} should produce an action.
		\item State vectors should come from elsewhere on the device: this may either be a dedicated island/core cluster (i.e., also performing INT management), or it may be a packet from the network.
	\end{itemize}
	\item Action Handling.
	\begin{itemize}
		\item A state vector is tile coded, action probabilities are produced, and an action is chosen according to current choice of $\epsilon$ (exploration parameter).
		\item \emph{Open question: how to configure action destination? This should either be another core (after an optional transformation), a packet sent out to a destination, maybe talk back to the controller.}
	\end{itemize}
	\item Learning.
	\begin{itemize}
		\item \emph{Open question: how to safely \& reasonably map state back to incoming rewards?}
		\item \emph{Open question: Can/should this also support on-chip reward signals?}
	\end{itemize}
\end{itemize}

\section{What's implemented 'til now?}
System?
\begin{itemize}
	\item Core-to-core communication.
	\item P4 parser for relevant packets.
	\item Packet copy + handoff from P4 to RL MEs.
	\item Dynamic base configuration.
	\item Dynamic tiling configuration.
	\item Policy installation via block-copy.
	\item Packet generation for above config functions, as well as policy quantisation and splitting.
	\item Tile-coding of quantised states on-chip, and state packet receipt.
\end{itemize}

Experiments?
\begin{itemize}
	\item Policy quantisation divergence \& accuracy (via mininet).
\end{itemize}

\section{Intermediate steps?}
System?
\begin{itemize}
	\item Action computation on-chip.
	\begin{itemize}
		\item \emph{\num{1} week.}
	\end{itemize}
	\item Figure out PIF rule installation.
	\begin{itemize}
		\item \emph{\num{1} week.}
		\item Rationale: understand and describe mechanism, do NOT implement. We know that we can save 1 controller RTT.
	\end{itemize}
	\item Setup other core as event source.
	\begin{itemize}
		\item \emph{\num{2} week.}
		\item Idea: proxy for idea of INT core mentioned elsewhere, combined with the main periodic event sending that the last RL work relied upon. So, say "you can put that state management and extraction on another core, here's what the comms cost is" or similar.
	\end{itemize}
	\item Implement rudimentary training.
	\begin{itemize}
		\item \emph{\num{2} week.}
	\end{itemize}
\end{itemize}

>>>>

@@

\DefineBibliographyStrings{english}{%
	andothers = {\emph{et al}\adddot}
}
\DeclareFieldFormat[inproceedings]{url}{}
\DeclareFieldFormat[article]{url}{}
%\DeclareFieldFormat[inproceedings]{doi}{}
%\DeclareFieldFormat[article]{doi}{}
%\DeclareFieldFormat[inproceedings]{editor}{}
%\DeclareFieldFormat[proceedings]{editor}{}
%\DeclareFieldFormat[article]{editor}{}

\newcommand{\fakepara}[1]{\noindent\textbf{#1:}}

@@

% DOI
\acmDOI{12345}

% ISBN
\acmISBN{67890}

%Conference
\acmConference[SOSR '21]{}
%\acmYear{2021}
%\copyrightyear{}

@@

%	\author{Paper \# XXX, XXX pages}
	 \author{Kyle A. Simpson}
	 \orcid{0000-0001-8068-9909}
	 \affiliation{%
	   \institution{University of Glasgow}
	   \city{Glasgow} 
	   \country{Scotland}
	 }
	 \email{k.simpson.1@research.gla.ac.uk}
	 \author{Dimitrios P. Pezaros}
\orcid{0000-0003-0939-378X}
	 \affiliation{%
	\institution{University of Glasgow}
	\city{Glasgow} 
	\country{Scotland}
}
\email{Dimitrios.Pezaros@gla.ac.uk}

@@

This paper contributes:
\begin{itemize}
	\item A thing (?? REFERENCE SECTION). Here's how that thing is utterly fantastic and will change the world, or not.
\end{itemize}

\section{Motivation}\label{sec:motivation}
\fakepara{Reinforcement learning in asynchronous environments}
RL techniques allow an \emph{agent} to learn how to iteratively interact with a system in an optimal fashion.
However, there is some degree of divergence between their theory and implementation.
Consider \cref{fig:state-slip}: the traditional formulation of a Markov decision process assumes that an agent receives a new view of the world's state at fixed time intervals, and then decides upon and executes an action instantly.
The reality is that state information takes time to traverse the network, service times are offset by how quickly hosts respond to interrupts and deserialise requests, and action preference lists are often computed via expensive policy approximations.
Action installation also incurs costs in fields such as network administration, initially to contact the controller and then for those actions to be installed via the control plane.

These delays (and variance thereof) add noise to the state-action mapping being learned, which has a potent reduction to learning rate and final accuracy, even for simple grid world tasks according to \textcite{DBLP:journals/firai/TravnikMSP18}.
They in turn show that reordering algorithmic steps can reduce these costs for online 1-step algorithms, but that reducing this further requires detailed agent-environment co-design.
This principle has influenced the design of real network use cases, such RL-based congestion-control algorithms~\parencite{DBLP:journals/corr/abs-1910-04054}, showing that asynchrony is necessary for high-speed applications.
Achieving this often requires that state measurements are combined or coalesced~\parencite{DBLP:journals/corr/abs-1910-04054,DBLP:journals/tnsm/SimpsonRP20} while expensive computations are ongoing.
`Stopping the world' in the algorithmic sense causes significant performance degradation, as inference takes up to \SI{30}{\milli\second} in the above work on congestion control, or any time-sensitive control problems.

\begin{figure}
	\begin{subfigure}{0.45\linewidth}
		\resizebox{\linewidth}{!}{
			\centering
			\begin{tikzpicture}
				\node[circle, draw] (state) {$S$};
				\node[circle, draw] (state') at ($(state) + (0, -2)$) {$S'$};

@@

				\draw[dotted, ->, bend left = 30] (state) -- (agent);
				\draw[->] (agent) -- (action);
				\draw[dotted, ->] (action) -- (state);
		\end{tikzpicture}}
		\caption{Theory: state measurement, action computation, and learning are zero-cost.}
	\end{subfigure}
	\begin{subfigure}{0.45\linewidth}
		\resizebox{0.75\linewidth}{!}{
			\centering
			\begin{tikzpicture}
				\node[circle, draw] (state) {$S$};
				\node[circle, draw] (state') at ($(state) + (0, -1.5)$) {$S'$};
				\node[circle, draw] (state'') at ($(state') + (0, -1.5)$) {$S''$};

@@

?? Find some cites citing the relevance of this problem wrt. self-driving cars, robotics, etc.

The solution we propose is to make use of the recent wave of programmable network devices to \emph{bring reinforcement learning to the data plane}---placing state measurement ($t_1$), low-cost decision-making processes ($t_2$), and controlled systems ($t_3$) as close to one another as possible.
In networks, actions are most likely to be installed on backbone switches, bump-in-the-wire NICs or middleboxes, and in the NICs of end-hosts.
Ideally, these functions which comprise an RL agent would all be collocated on the same chip or device, but this is easier said than done. 
Both programmable devices and the network environment make this more difficult, as we'll examine in the sequel.

\fakepara{Programmable hardware capabilities}
The introduction of the P4~\parencite{DBLP:journals/ccr/BosshartDGIMRSTVVW14} language has led to explosive growth in the research community surrounding in-network computation and offloading.
Providing a single language which is compatible with network devices of many form factors has been instrumental in the development of novel fine-grained traffic measurement approaches~\parencite{DBLP:conf/sigcomm/GuptaHCFRW18,DBLP:conf/sigcomm/ChenFKRR18,DBLP:conf/sosr/GhasemiBR17}.
However, this requires that there be a reasonably consistent interface and behaviour between device classes.
This need for underpins the \emph{Programmable Switch Architecture} (PSA)~\parencite{p4-psa}, which defines a conceptual model of match-action tables divided into ingress and egress pipelines.
The PSA presents a sensible lower bound on the device capabilities required to implement a P4 dataplane, but the reality is more complex and interesting.

We would be remiss to believe that the P4 and the PSA define all capabilities that programmable hardware can support.
Many compatible devices have legacies predating these developments, for instance many-core SOC-based Netronome SmartNIC~\parencite{netronome-smartnic}, and NetFPGA SUME~\parencite{DBLP:journals/micro/ZilbermanACM14} NICs.
The first of these compiles from P4$\rightarrow$Micro-C$\rightarrow$bytecode (with Micro-C externs), while the latter can combine the P4$\rightarrow$NetFPGA toolchain~\parencite{DBLP:conf/fpga/IbanezBMZ19} with arbitrary circuit externs---both exposing further device capabilities beyond the specification.
Even Intel Tofino~\parencite{barefoot-intel} ASICs, which architecturally mirror the PSA, expose additional matching capabilities and ALU functions via the Tofino Native Architecture.
Regardless, floating-point operations key to ML/RL workloads are very rarely supported outside of DL accelerators (which, in turn, lack packet switching functionality).
Moreover, the limited memory/block RAM and per-packet timing constraints endemic to these devices make this class of in-NIC offloading challenging.

Luckily, the task of bringing machine learning models to resource-constrained environments is well-studied.
Quantisation and alternative data formats have been suggested to make ML inference feasible on resource- and power-limited platforms, work around hardware constraints, or compute faster and more efficiently.
Lower bit-depths reduce memory footprints, and improve throughput in designs such as \emph{bfloat16}~\parencite{bfloat16-blog} in Google TPUs~\parencite{DBLP:journals/sigops/XieDMKVZT18}, \emph{hfp8}~\parencite{DBLP:conf/nips/SunCCWVSCZG19}, and other floating point formats~\parencite{DBLP:journals/corr/abs-2007-01530}.
In many cases, accuracy losses for doing so are negligible.
Much of this work goes further still towards integer~\parencite{tensorrt-8bit} or binarised~\parencite{DBLP:journals/corr/MiyashitaLM16,DBLP:conf/eccv/RastegariORF16,DBLP:journals/corr/KimS16,DBLP:conf/nips/HubaraCSEB16} representations, sacrificing dynamic range for simpler arithmetic operations.
This has the important side effect of making ML policy inference tractable on hardware without floating-point capabilities, which other works have leveraged to bring neural networks to the dataplane~\parencite{DBLP:journals/corr/abs-2009-02353,DBLP:conf/sigcomm/SanvitoSB18,DBLP:journals/corr/abs-1801-05731}.

To make RL workloads feasible under such constraints, we propose the use of quantised, fixed-point representations such as $Qm.n$, which allow us to evaluate and update policies using only integer arithmetic.
This is not only essential in performing this work, but also serves as a mechanism for reducing the processing and memory costs of function approximation ($t_2$, \cref{fig:state-slip}).
Crucially, we train our focus on devices with a low port density, such as SmartNICs and NetFPGA devices.
The designs of these devices make it reasonable to move policy processing \emph{outside of the packet pipeline}; SmartNICs expose many additional programmable cores, and NetFPGAs allow for the synthesis of independent functional units.
The main benefit of doing so is that core control logic can be moved as close to the device as possible \emph{without impacting packet processing rates}.
When combined with the above P4-driven techniques for in-network flow/device measurement, we at last have the mechanisms to collocate the key processes of an RL agent.
Moreover, the base P4 dataplane can be used to simplify parsing logic and offer runtime control over which flows/packets are monitored, alongside other packet actions.
This again fits our goals of integrating RL techniques directly within bump-in-the-wire installations and at end-hosts.

?? relate update/execute operations here? i.e., policy execute needs only ALU add, train needs add+mul+shr. or in RL section? Should that come first?

\fakepara{Online training}
Some problems either evolve over time in unpredictable ways, or cannot be easily modelled.
This can then make \emph{online} learning of the task an attractive prospect, using a single available stream of experience.
DNNs, particularly when used as the basis for an RL agent's policy, require vast amounts of experience to converge on an accurate parameter set.
In many problem domains, this equates to training from compute-years worth of distributed offline simulations, which is at odds with the need to adapt to changes in the underlying problem \emph{as they happen}.
Achieving stable learning requires sizeable batches of gradients to be computed, potentially using the entire experience replay from each simulation.
Still, the computational cost of gradient computation via backpropagation is significant on embedded hardware.

To then achieve \emph{online} learning, we return our focus to older \emph{classical} reinforcement learning methods.
In particular, we suggest tile-coding~\cite[pp. \numrange{217}{221}]{RL2E}, with one-step temporal-difference learning algorithms such as Sarsa or Q-learning~\cite[pp. \numrange{129}{132}]{RL2E}.
These choices have important benefits for in-NIC execution.
Simpler function approximators do not require batches of inputs to learn in a more stable way, negating the additional memory required to store experience replays.
Tile-coding in particular admits many optimisations, being an embarrassingly parallel problem.
Furthermore, gradient calculation is trivial for linear methods like tile coding, particularly when compared to the computational cost of backpropagation required by neural networks.
Finally, the choice of single-step algorithms (as opposed to $n$-step or Monte Carlo methods) bounds the amount of per-trace state required for online learning to just the last state-action pair.

?? Probably want some cites on the need for batching in NN methods, even though this is understood. DL book?

\fakepara{To-be-organised}
Interesting thing I realised from \textcite{DBLP:conf/sigcomm/NeugebauerAZAL018}: GPU and NIC might cause IOMMU contention? Making line-rate offload of DDN to GPU impractical. Unsure if I want to twin this with later measurements/plots.

\section{Background}

\subsection{Classical Reinforcement Learning}
?? Move into background parent section?

?? RL~\parencite{RL2E}

?? Use this to explain differences between the modern state of the art and older works.

?? Probably give a rough introduction to the algorithm, show how computation needs only addition---particularly when using Qm.n fixed point.

\subsection{Hmm..}
Should motivation be in here? Unsure what other preliminaries need to be explained.

\section{Design and Implementation}\label{sec:design}
Based on the design principles, problems, and potential solutions outlined throughout \cref{sec:motivation}, we present our design for an in-NIC, task-independent, online reinforcement learning system.
\Cref{fig:netro-arch} outlines our design and implementation on Netronome SmartNIC hardware in pursuit of this goal.
At a high level, ...

\subsection{Policy Storage}
?? Blocks of dense quantised tiles: each an action preference list.

?? Tiling data is split into regions of memory based on access time: higher-dimension tilings are sent to further NUMA regions (EMEM/IMEM).

\subsection{Action Computation}
?? Split across MEs.

?? each ME can then subdivide by factor of 4 among contexts.

\subsection{Communication}
?? Free-list of packet handles shared between P4 pipeline threads for state/config/action handover.

?? multi-producer-multi-consumer -- rings used to signal cores, place work, etc., combined with above free-list

?? used for both directions of communication.

\subsection{Notes from PGR}

The main gist of this design is that RL operations \emph{do not happen on the core data path}, which ensures that action computation, installation, and updates don't stall packet transmission or affect throughput.
In the default deployment of a P4 packet processing pipeline, several of these cores go unused, making this paradigm possible.
These dedicated cores (\emph{micro-engines} grouped by \emph{island} in NFP terminology) are then responsible for processing requests, computing actions, and updating the underlying policy in real time.
The main interaction model is that platform-specific IPC mechanisms (message rings) are used to \emph{push} configuration, state vectors, and reward measurements to the RL system.
These same mechanisms are used by other cores on the same device to \emph{pull} output actions from the RL system.
Both input and output can occur on any other core of the device, i.e., as part of P4 \texttt{extern} plugins, while the P4 control plane itself provides granular control over which flows are monitored or affected.
An input state vector \emph{always} induces an action, and if desired updates the policy using either an included reward or one retrieved from memory according to a (configurable) key in the state.
Runtime reconfiguration and interaction occur via the control and/or dataplane: the ease of use of the P4 pipeline's match and custom protocol parsers, combined with the dedicated input pipe to the NIC's controller (the host machine), allow these to be cleanly separated or combined as needed.

?? Expand: this is based on PGR Year 3 atm.

\subsection{Targetting other device classes}
While this design caters primarily to Netronome's SmartNIC devices, many aspects of this design have clear analogues in NICs of similar form-factor, such as their close NetFPGA SUME cousins.
For instance, the availability of other cores can be handled by designing additional off-path functional units, much like the floating-point adders used by \textcite{DBLP:conf/isca/LiLYCSH19}.
Rather than general-purpose cross-core communication, an RL functional unit would be hard-wired to and from the custom actions that interface with it, using the P4$\rightarrow$NetFPGA toolchain to offer the base framework and granular control over packet and flow selection.

?? Fill out above with more specific instances.

Bringing this to high-port density devices such as Intel Tofino-powered switches is more difficult.
As discussed in \cref{sec:motivation}, Tofino ASICs map closely to the P4 PSA, meaning that there are no spare asynchronous general-purpose compute units to place this work on.
A potential solution is to divide RL actions across several received packets (i.e., iteratively computing a portion of the action preference list each time) until any further work would delay outbound transmission.
This, however, would introduce new issues surrounding concurrent accesses, work splitting, and altered timescales for learning: we leave their treatment and examination to future work.

@@

\begin{figure}
	\centering
	\resizebox{1.2\linewidth}{!}{\begin{tikzpicture}
			\node (sw2) {};

			\node[fill=white!80!uofgcobalt, draw=black, minimum height=5.5cm, minimum width=2.9cm, below right= 0.1 of sw2.north west](p4-master){};
			\node[below right, inner sep=2pt] at (p4-master.north west) {\small P4 Datapath};
			\node[fill=white!90!uofgcobalt, draw, rectangle, rounded corners=0.05cm, below=0.8] (p4-rules) at (p4-master.north) {\begin{varwidth}{2.5 cm}\scriptsize \centering Flow Control Rules\end{varwidth}};
			\node[fill=white!90!uofgcobalt, draw, rectangle, rounded corners=0.05cm, below=0.1] (n1) at (p4-rules.south) {\begin{varwidth}{2.5 cm}\scriptsize \centering Extern-based Action\end{varwidth}};
			\node[fill=white!90!uofgcobalt, draw, rectangle, rounded corners=0.05cm, below=0.1] (n2) at (n1.south) {\begin{varwidth}{2.5 cm}\scriptsize \centering ...\end{varwidth}};
			\node[fill=white!90!uofgcobalt, draw, rectangle, rounded corners=0.05cm, below=0.1] (oswlc) at (n2.south) {\begin{varwidth}{1.5 cm}\scriptsize \centering Control Packet Match\end{varwidth}};
			\node[fill=white!90!uofgcobalt, draw, rectangle, rounded corners=0.05cm, below=0.1] (n3) at (oswlc.south) {\begin{varwidth}{2.5 cm}\scriptsize \centering ...\end{varwidth}};
			\node[fill=white!90!uofgcobalt, draw, rectangle, rounded corners=0.05cm, below=0.1] (n4) at (n3.south) {\begin{varwidth}{2.5 cm}\scriptsize \centering Network Telemetry\end{varwidth}};

			\node (pkt-text) at ($(p4-master.north) + (0.0,0.6)$) {\begin{varwidth}{2.5 cm}\scriptsize \centering Packet In\end{varwidth}};

			\node (pkt-text-out) at ($(p4-master.south) + (0.0,-0.6)$) {\begin{varwidth}{2.5 cm}\scriptsize \centering Packet Out\end{varwidth}};

			\node[fill=white!80!uofgsunshine, draw=black, minimum height=5cm, minimum width=5cm](rli) at ($(p4-master.east) + (5, -1)$){};
			\node[below right, inner sep=2pt] at (rli.north west) (rli-text) {\small RL Island};

			\node[draw=black, dotted, minimum height=1cm, minimum width=2cm](wq-out) at ($(rli-text.south) + (0.5,-0.5)$){};
			\node[below right, inner sep=2pt] at (wq-out.north west) {\small EMEM};
			\node[fill=white!90!uofgsunshine, draw, rectangle, rounded corners=0.05cm, above=0.1] (wq-in) at (wq-out.south) {\begin{varwidth}{2.5 cm}\scriptsize \centering Input Queue\end{varwidth}};

@@

			\draw[->] (oswlc.east)  to node[midway, above, sloped] {\tiny Control, Requests} (wq-in.west);
			\draw[->] (n4.east)  to node[midway, above, sloped] {\tiny State} (wq-in.west);
			\draw[<->, dotted] (wq-in.east)  to (loop.west);
			\draw[->, dotted] (loop.south) to (policy-cls-out);

			\draw[->] (loop.north)  to node[midway, above, sloped] {\tiny Actions} (wq-actin.south);

			\draw[->] (wq-actin.west)  to node[midway, above, sloped] {\tiny Actions} (n1.east);

@@

			\draw[->] (pkt-text)  to  (p4-master.north);
			\draw[->] (p4-master.south)  to  (pkt-text-out);

			\draw[->] (loop) to node[midway, above, sloped] {\tiny NN} (po-me-1);
			\draw[->] (po-me-1) to node[midway, above, sloped] {\tiny NN} (po-me-2);
			\draw[->] (po-me-2) to node[midway, above, sloped] {\tiny NN} (po-me-3);
			\draw[->] (po-me-3) to [out=-45,in=0,looseness=2] node[midway, above, sloped] {\tiny NN} (loop);
	\end{tikzpicture}}
	\caption{Architecture for generalised RL agent on Netronome hardware.\label{fig:netro-arch}}
\end{figure}

@@

			\item[$t_3$]
			\begin{description}
				\item[PDP] Modify local tables across ME/Island boundaries (same device), verify rule present (or get ACK).
				\item[VNF] Generate OpenFlow control messages, send OF message to switch, verify rule present (or get ACK).
			\end{description}
		\end{description}
		\item HYPOTHESIS: PDP will present a reduction in these three times for per-flow state.
		\item QUESTION: Does same apply to learning?
		\item NOTE: since we're not actually assessing the \emph{quality} of output decisions, we can cut corners on what data is sent out and where. I.e., in-band network telemetry island can be skipped, if we move over representative volumes of state to/from the RL island with approximately correct delay as recorded by other studies.
	\end{enumerate}
\end{itemize}
I can think of one way to evaluate this if we employ classification performance:
\begin{enumerate}
	\item Run the existing work~\parencite{DBLP:journals/tnsm/SimpsonRP20}, acquire a policy for one node after full info sharing.
	\item Load this policy onto the RL Island (\cref{fig:netro-arch}).
	\item Either feed in raw state packets, or feed in traffic if we can get INT working.
	\item For each level of quantisation needed, examine average performance.
	\item Don't compare against old numbers for final perf: use a ``VNF'' setup similar to above---this will probably suffer a bit compared to old results, where everything was on the same machine. Make it relative/normalised.
\end{enumerate}

\section{Bullets on (intended) design.}
See also: \cref{fig:netro-arch}.
\begin{itemize}
	\item High-level overview.
	\begin{itemize}
		\item RL operations \emph{do not happen on the core path}.
		\item Another island (or set of cores) is dedicated to handling inputs, configuration, computation.
		\item Cross-island communication happens over \emph{rings} (Netronome-specific).
		\begin{itemize}
			\item E.g., this is how the P4 firmware propagates rules from master island to P4 workers.
			\item My impl.\ specific---shared freelist for memory pointers in EMEM.
		\end{itemize}
		\item Controller is \emph{always} co-hosted, interfaces over PCI-e. This also applies to Barefoot Tofino.
	\end{itemize}
	\item Policy configuration.
	\begin{itemize}
		\item (Re)-configurable at runtime.
		\item Custom DSCP marker acts as indicator of RL packets (i.e., bother with parsing past UDP).
		\item P4 parser extracts headers, parser state and payload passed over ring to another core dedicated to RL operations.
		\item Policy dimensions, size, action count, tiling strategies, learning parameters ($\alpha, \gamma, \epsilon$, related falloffs) can all be chosen and changed live.
	\end{itemize}
	\item State Handling.
	\begin{itemize}
		\item State input to model \emph{always} should produce an action.
		\item State vectors should come from elsewhere on the device: this may either be a dedicated island/core cluster (i.e., also performing INT management), or it may be a packet from the network.
	\end{itemize}
	\item Action Handling.
	\begin{itemize}
		\item A state vector is tile coded, action probabilities are produced, and an action is chosen according to current choice of $\epsilon$ (exploration parameter).
		\item \emph{Open question: how to configure action destination? This should either be another core (after an optional transformation), a packet sent out to a destination, maybe talk back to the controller.}
	\end{itemize}
	\item Learning.
	\begin{itemize}
		\item \emph{Open question: how to safely \& reasonably map state back to incoming rewards?}
		\item \emph{Open question: Can/should this also support on-chip reward signals?}
	\end{itemize}
\end{itemize}

\section{What's implemented 'til now?}
System?
\begin{itemize}
	\item Core-to-core communication.
	\item P4 parser for relevant packets.
	\item Packet copy + handoff from P4 to RL MEs.
	\item Dynamic base configuration.
	\item Dynamic tiling configuration.
	\item Policy installation via block-copy.
	\item Packet generation for above config functions, as well as policy quantisation and splitting.
	\item Tile-coding of quantised states on-chip, and state packet receipt.
	\item Action computation on-chip.
	\item Rudimentary, single-core training.
\end{itemize}

Experiments?
\begin{itemize}
	\item Policy quantisation divergence \& accuracy (via mininet).
	\item Measure DPDK SmartNIC RTT as a lower bound on controller access time.
	\begin{itemize}
		\item Same mechanism applies to other manufacturers (Barefoot), so generalisable.
		\item My results are non-DPDK, but there is a good study~\parencite{DBLP:conf/sigcomm/NeugebauerAZAL018} on exactly this.
	\end{itemize}
	\item Measure basic on-switch times: ring-ring communication, tile-coding cost, action compute cost, learning cost.
\end{itemize}

Known/understood (but not implemented)?
\begin{itemize}
	\item Figure out PIF rule installation.
	\begin{itemize}
		\item Not doable through P4 na\"{i}vely. Handing back rules to controller is expensive: need another action to deliberately use its own hash table etc., then batch back as faster P4 rules.
	\end{itemize}
\end{itemize}
